Hrvatska u kontekstu europskih konvergencijskih klubova

### 2024-09-25 Paper.Rmd
cool

```{r fracr1 ,  echo=FALSE, message=FALSE, warning=FALSE,  fig.width=10, fig.height=6}

apply_robinson_estimator <- function(data, group_name) {
  result <- arfima::arfima(data$diff_log_gdp_per_capita)

  # Extract the fractional differencing parameter (dfrac)
  d.f <- result$modes[[1]]$dfrac
  se.d.f <- result$modes[[1]]$se[1]  # Assuming the standard error for dfrac is the first element in the se vector

  # Check if the standard error is available
  if (is.null(se.d.f)) {
    se.d.f <- NA
    p.value <- NA
  } else {
    # Calculate the t-statistic for d.f
    t.stat <- d.f / se.d.f

    # Calculate the p-value for the t-statistic (two-tailed test)
    p.value <- 2 * pt(-abs(t.stat), df = length(data$diff_log_gdp_per_capita) - 1)
  }

  tibble(
    group_name = group_name,
    robinson_d = d.f,
    p_value = p.value
  )
}

# Calculate differences for each group
diff_croatia_vs_nms8 <- calculate_differences(gdp_croatia, gdp_nms8)
diff_croatia_vs_nms12 <- calculate_differences(gdp_croatia, gdp_nms12)
diff_croatia_vs_se4 <- calculate_differences(gdp_croatia, gdp_se4)
diff_croatia_vs_eu15 <- calculate_differences(gdp_croatia, gdp_eu15)

# Apply Robinson estimator to each group
robinson_croatia_vs_nms8 <- apply_robinson_estimator(diff_croatia_vs_nms8, "Croatia vs NMS8")
robinson_croatia_vs_nms12 <- apply_robinson_estimator(diff_croatia_vs_nms12, "Croatia vs NMS12")
robinson_croatia_vs_se4 <- apply_robinson_estimator(diff_croatia_vs_se4, "Croatia vs SE4")
robinson_croatia_vs_eu15 <- apply_robinson_estimator(diff_croatia_vs_eu15, "Croatia vs EU15")

# Combine Robinson test results
robinson_results <- bind_rows(robinson_croatia_vs_nms8, robinson_croatia_vs_nms12, robinson_croatia_vs_se4, robinson_croatia_vs_eu15)

# Print the Robinson test results
kable(
  robinson_results, 
  caption = "Robinson Estimator Results for Income Convergence",
  col.names = c("Group Name", "Robinson d", "P-value"),
  digits = 4,
  align = 'c'
) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
add_header_above(c(" " = 1, "Estimation Results" = 2)) # Adds a header above the columns

@@

```{r}
clean_data <- na.omit(diff_croatia_vs_nms8$diff_log_gdp_per_capita)
result <- fracdiff(clean_data)
print(result$d)
```

```{r eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}

library(rugarch)
library(fracdiff)
library(arfima)
library(longmemo)

# Example series (use your actual differenced series)
diff_series_list <- list(
  "Croatia vs NMS8" = na.omit(diff_croatia_vs_nms8$diff_log_gdp_per_capita),
  "Croatia vs NMS12" = na.omit(diff_croatia_vs_nms12$diff_log_gdp_per_capita),
  "Croatia vs SE4" = na.omit(diff_croatia_vs_se4$diff_log_gdp_per_capita),
  "Croatia vs EU15" = na.omit(diff_croatia_vs_eu15$diff_log_gdp_per_capita)
)

# Initialize result storage
results <- list()

# Loop through each differenced series
for (group_name in names(diff_series_list)) {
  series <- diff_series_list[[group_name]]

  # Remove or handle missing values and ensure it's a time series
  clean_series <- ts(na.omit(series))  # Ensure no NaNs and convert to time series

  # Store results for this group
  result_list <- list()

  # Apply fracdiff()
  fracdiff_result <- fracdiff(clean_series)
  result_list$fracdiff_d <- fracdiff_result$d

  # Apply arfima() from arfima package
  arfima_fit <- arfima(clean_series)
  result_list$arfima_d <- arfima_fit$modes[[1]]$dfrac

  # Apply autoarfima() from rugarch package (simplified)
  # IMPORTANT: Simplified function call for autoarfima
  # autoarfima_fit <- autoarfima(clean_series)  # Removed the 'arfima = TRUE' argument
  # result_list$autoarfima_d <- coef(autoarfima_fit)["d"]

  # Apply GPHfit() from longmemo package (Geweke-Porter-Hudak)
  gph_result <- GPHfit(clean_series)
  result_list$gph_d <- gph_result$d

  # Store results for this group
  results[[group_name]] <- result_list
}

# Display the results
results_df <- do.call(rbind, lapply(names(results), function(group) {
  cbind(group, as.data.frame(t(results[[group]])))
}))

print(results_df)

```{r fracr2 ,  echo=FALSE, message=FALSE, warning=FALSE,  fig.width=10, fig.height=6}

# GPH estimator
gph_test_results <- function(data, group_name) {
  result <- fracdiff::fdGPH(data$diff_log_gdp_per_capita)

  # Return a tibble with group name and estimated d value
  tibble(
    group_name = group_name,
    gph_d = result$d
  )
}

gph_croatia_vs_nms8 <- gph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8")
gph_croatia_vs_nms12 <- gph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12")
gph_croatia_vs_se4 <- gph_test_results(diff_croatia_vs_se4, "Croatia vs SE4")
gph_croatia_vs_eu15 <- gph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15")

# Combine GPH test results
gph_results <- bind_rows(gph_croatia_vs_nms8, gph_croatia_vs_nms12, gph_croatia_vs_se4, gph_croatia_vs_eu15)

# Print the GPH test results
kable(
  gph_results, 
  caption = "GPH Estimator Results for Income Convergence",
  col.names = c("Group Name", "GPH d"),
  digits = 4,
  align = 'c'
) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
add_header_above(c(" " = 1, "Estimation Results" = 1))  # Adds a header above the columns

```{r fracr3 ,  echo=FALSE, message=FALSE, warning=FALSE,  fig.width=10, fig.height=6}
# Define the P-GPH estimator function
pgph_test_results <- function(data, group_name) {
  # Run the Phillips-Perron test (as a proxy for P-GPH)
  result <- ur.pp(data$diff_log_gdp_per_capita, type = "Z-alpha", model = "trend")

  # Extract the test statistic
  test_stat <- result@teststat[1]  # Z-alpha statistic

  # Check the structure of the cval object
  critical_values <- result@cval

  # Safely extract critical values
  critical_1_percent <- ifelse(nrow(critical_values) >= 1, critical_values[1, 1], NA)
  critical_5_percent <- ifelse(nrow(critical_values) >= 2, critical_values[2, 1], NA)
  critical_10_percent <- ifelse(nrow(critical_values) >= 3, critical_values[3, 1], NA)

  # Return as a data frame (or tibble)
  tibble(
    group_name = group_name,
    pgph_stat = test_stat,
    critical_1_percent = critical_1_percent,
    critical_5_percent = critical_5_percent,
    critical_10_percent = critical_10_percent
  )
}

# Apply the P-GPH estimator to each group
pgph_croatia_vs_nms8 <- pgph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8")
pgph_croatia_vs_nms12 <- pgph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12")
pgph_croatia_vs_se4 <- pgph_test_results(diff_croatia_vs_se4, "Croatia vs SE4")
pgph_croatia_vs_eu15 <- pgph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15")

# Combine P-GPH test results
pgph_results <- bind_rows(pgph_croatia_vs_nms8, pgph_croatia_vs_nms12, pgph_croatia_vs_se4, pgph_croatia_vs_eu15)

# Print the P-GPH test results

kable(
  pgph_results, 
  caption = "P-GPH Test Results for Income Convergence",
  col.names = c("Group Name", "P-GPH Statistic", "Critical 1%", "Critical 5%", "Critical 10%"),
  digits = 4,
  align = 'c'
) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
add_header_above(c(" " = 1, "Estimation Results" = 4))  # Adds a header above the columns

>>>>

```{r fracr2 ,  echo=FALSE, message=FALSE, warning=FALSE,  fig.width=10, fig.height=6}
# Function to apply different estimators (fdGPH, fdSperio, fracdiff)
gph_test_results <- function(data, group_name, bandwidth, estimator) {
  if (estimator == "fdGPH") {
    # Apply the GPH estimator with a specified bandwidth
    result <- fracdiff::fdGPH(data$diff_log_gdp_per_capita, bandw.exp = bandwidth)
    d_value <- result$d
  } else if (estimator == "fdSperio") {
    # Apply the Sperio estimator
    result <- fracdiff::fdSperio(data$diff_log_gdp_per_capita)
    d_value <- result$d
  } else if (estimator == "fracdiff") {
    # Apply the fracdiff MLE estimator
    result <- fracdiff::fracdiff(data$diff_log_gdp_per_capita)
    d_value <- result$d
  }

  # Return a tibble with group name, estimated d value, and bandwidth or method used
  tibble(
    group_name = group_name,
    gph_d = d_value,
    bandwidth = bandwidth
  )
}

# Define bandwidths and methods
bandwidths <- c(0.4 , 0.5, 0.6, 0.7, 0.8, 0.9)
methods <- c("fdGPH", "fdSperio", "fracdiff")

# For each estimator, apply the test to different groups and bandwidths

# GPH Estimator (fdGPH)
gph_croatia_vs_nms8_fdGPH <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8", .x, "fdGPH"))
gph_croatia_vs_nms12_fdGPH <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12", .x, "fdGPH"))
gph_croatia_vs_se4_fdGPH <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_se4, "Croatia vs SE4", .x, "fdGPH"))
gph_croatia_vs_eu15_fdGPH <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15", .x, "fdGPH"))

gph_results_long_fdGPH <- bind_rows(gph_croatia_vs_nms8_fdGPH, gph_croatia_vs_nms12_fdGPH, gph_croatia_vs_se4_fdGPH, gph_croatia_vs_eu15_fdGPH)
gph_results_wide_fdGPH <- gph_results_long_fdGPH %>%
  tidyr::pivot_wider(names_from = bandwidth, values_from = gph_d, names_prefix = "Bandwidth_")

# Sperio Estimator (fdSperio)
gph_croatia_vs_nms8_fdSperio <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8", .x, "fdSperio"))
gph_croatia_vs_nms12_fdSperio <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12", .x, "fdSperio"))
gph_croatia_vs_se4_fdSperio <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_se4, "Croatia vs SE4", .x, "fdSperio"))
gph_croatia_vs_eu15_fdSperio <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15", .x, "fdSperio"))

gph_results_long_fdSperio <- bind_rows(gph_croatia_vs_nms8_fdSperio, gph_croatia_vs_nms12_fdSperio, gph_croatia_vs_se4_fdSperio, gph_croatia_vs_eu15_fdSperio)
gph_results_wide_fdSperio <- gph_results_long_fdSperio %>%
  tidyr::pivot_wider(names_from = bandwidth, values_from = gph_d, names_prefix = "Bandwidth_")

# Fracdiff Estimator (fracdiff)
gph_croatia_vs_nms8_fracdiff <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8", .x, "fracdiff"))
gph_croatia_vs_nms12_fracdiff <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12", .x, "fracdiff"))
gph_croatia_vs_se4_fracdiff <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_se4, "Croatia vs SE4", .x, "fracdiff"))
gph_croatia_vs_eu15_fracdiff <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15", .x, "fracdiff"))

gph_results_long_fracdiff <- bind_rows(gph_croatia_vs_nms8_fracdiff, gph_croatia_vs_nms12_fracdiff, gph_croatia_vs_se4_fracdiff, gph_croatia_vs_eu15_fracdiff)
gph_results_wide_fracdiff <- gph_results_long_fracdiff %>%
  tidyr::pivot_wider(names_from = bandwidth, values_from = gph_d, names_prefix = "Bandwidth_")

# Create and display tables for each estimator

# GPH Estimator Table
kable(
  gph_results_wide_fdGPH, 
  caption = "GPH Estimator (fdGPH) Results for Income Convergence by Bandwidth",
  col.names = c("Group Name", paste0("GPH d (", bandwidths, ")")),
  digits = 4,
  align = 'c'
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  add_header_above(c(" " = 1, "Estimation Results by Bandwidth" = length(bandwidths)))

# Sperio Estimator Table
kable(
  gph_results_wide_fdSperio, 
  caption = "GPH Estimator (fdSperio) Results for Income Convergence by Bandwidth",
  col.names = c("Group Name", paste0("GPH d (", bandwidths, ")")),
  digits = 4,
  align = 'c'
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  add_header_above(c(" " = 1, "Estimation Results by Bandwidth" = length(bandwidths)))

@@

```{r}
# Function to add significance stars based on p-value
add_significance_stars <- function(p_value) {
  if (p_value <= 0.001) {
    return("***")
  } else if (p_value <= 0.01) {
    return("**")
  } else if (p_value <= 0.05) {
    return("*")
  } else if (p_value <= 0.1) {
    return(".")
  } else {
    return("")
  }
}

# Function to apply different estimators (fdGPH, fdSperio) and calculate significance
gph_test_results <- function(data, group_name, bandwidth, estimator) {
  if (estimator == "fdGPH") {
    # Apply the GPH estimator with a specified bandwidth
    result <- fracdiff::fdGPH(data$diff_log_gdp_per_capita, bandw.exp = bandwidth)
    d_value <- result$d
  } else if (estimator == "fdSperio") {
    # Apply the Sperio estimator
    result <- fracdiff::fdSperio(data$diff_log_gdp_per_capita)
    d_value <- result$d
  }

  # Calculate the number of observations (n)
  n <- length(data$diff_log_gdp_per_capita)

  # Asymptotic standard error of the estimated d value
  se_d <- 1 / sqrt(n)

  # Z-score and p-value for significance testing
  z_score <- d_value / se_d
  p_value <- 2 * (1 - pnorm(abs(z_score)))  # Two-tailed p-value

  # Add significance stars
  significance <- add_significance_stars(p_value)

  # Return a tibble with group name, estimated d value, significance, and bandwidth
  tibble(
    group_name = group_name,
    gph_d = paste0(round(d_value, 4), significance),  # Append significance stars to d
    bandwidth = bandwidth
  )
}

# Define bandwidths and methods
bandwidths <- c(0.4 , 0.5, 0.6, 0.7, 0.8, 0.9)
methods <- c("fdGPH", "fdSperio")

# Apply the test to different groups and bandwidths

# GPH Estimator (fdGPH)
gph_croatia_vs_nms8_fdGPH <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8", .x, "fdGPH"))
gph_croatia_vs_nms12_fdGPH <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12", .x, "fdGPH"))
gph_croatia_vs_se4_fdGPH <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_se4, "Croatia vs SE4", .x, "fdGPH"))
gph_croatia_vs_eu15_fdGPH <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15", .x, "fdGPH"))

gph_results_long_fdGPH <- bind_rows(gph_croatia_vs_nms8_fdGPH, gph_croatia_vs_nms12_fdGPH, gph_croatia_vs_se4_fdGPH, gph_croatia_vs_eu15_fdGPH)
gph_results_wide_fdGPH <- gph_results_long_fdGPH %>%
  tidyr::pivot_wider(names_from = bandwidth, values_from = gph_d, names_prefix = "Bandwidth_")

# Sperio Estimator (fdSperio)
gph_croatia_vs_nms8_fdSperio <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8", .x, "fdSperio"))
gph_croatia_vs_nms12_fdSperio <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12", .x, "fdSperio"))
gph_croatia_vs_se4_fdSperio <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_se4, "Croatia vs SE4", .x, "fdSperio"))
gph_croatia_vs_eu15_fdSperio <- purrr::map_df(bandwidths, ~ gph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15", .x, "fdSperio"))

gph_results_long_fdSperio <- bind_rows(gph_croatia_vs_nms8_fdSperio, gph_croatia_vs_nms12_fdSperio, gph_croatia_vs_se4_fdSperio, gph_croatia_vs_eu15_fdSperio)
gph_results_wide_fdSperio <- gph_results_long_fdSperio %>%
  tidyr::pivot_wider(names_from = bandwidth, values_from = gph_d, names_prefix = "Bandwidth_")

# Create and display tables for each estimator

# GPH Estimator Table
kable(
  gph_results_wide_fdGPH, 
  caption = "GPH Estimator (fdGPH) Results for Income Convergence by Bandwidth with Significance",
  col.names = c("Group Name", paste0("GPH d (", bandwidths, ")")),
  digits = 4,
  align = 'c'
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  add_header_above(c(" " = 1, "Estimation Results by Bandwidth" = length(bandwidths)))

# Sperio Estimator Table
kable(
  gph_results_wide_fdSperio, 
  caption = "GPH Estimator (fdSperio) Results for Income Convergence by Bandwidth with Significance",
  col.names = c("Group Name", paste0("GPH d (", bandwidths, ")")),
  digits = 4,
  align = 'c'
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  add_header_above(c(" " = 1, "Estimation Results by Bandwidth" = length(bandwidths)))
